---
title: "In-class exercise 2"
author: "LIU Yijun, Edrian"
date: "2023-10-05"
output: pdf_document
editor_options: 
  chunk_output_type: inline
---

## 1: Codebook lookup

### 1.1 Indicators regarding the quality of education

In P381 of the Codebook, we can see two indicators regarding education:

1.  Education 15+ (e_peaveduc), which describes the average years of education among citizens older than 15.
2.  Educational inequality, Gini (e_peedgini), which describes the level of inequality of education achieved by the population aged 15 or older.

### 1.2 Data's coverage

For Education 15+: worldwide coverage, 1820-2022

For Educational inequality, Gini: worldwide coverage, 1850-2010

### 1.3 Sources

Clio Infra (clio-infra.eu)

## 2: Subset by columns

### 2.1 Country-year identifiers and indicators of education quality

Step 1: Loading library and data

```{r}
library(tidyverse)

d <- read_csv("data/vdem/1984_2022/vdem_1984_2022_external.csv")
```

Step 2: Selecting the identifiers and indicators

```{r}
d_edu <- d |>
  select(country_name, country_id, year, e_peaveduc, e_peedgini)
```

### 2.2 Renaming Columns

```{r}
d_edu <- d_edu |>
  rename("Country" = "country_name",
         "ID" = "country_id",
         "Year" = "year", 
         "Education_Level" = "e_peaveduc", 
         "Education_Inequality" = "e_peedgini")

glimpse(d_edu)
```

## 3. Subset by rows

### 3.1 Listing 5 country-years with highest education level

```{r}
d_edu |>
  slice_max(order_by = Education_Level, n = 5)
```

As the United Kingdom has 13 years of identical data (13.3 years), they are all shown in the result.

### 3.2 Listing 5 country-years with highest education inequality

```{r}
d_edu |>
  slice_max(order_by = Education_Inequality, n = 5)
```

## 4. Summarize the data

### 4.1 Check data availability

#### 4.2.1 Data missing for each country

For `Education_Level`:

```{r}
d_edu |>
  mutate(edu_level_missing_country = as.numeric(is.na(Education_Level))) |>
  group_by(Country) |>
  summarize(num_edu_level_missing_country = sum(edu_level_missing_country))
```

For `Education_Inequality`:

```{r}
d_edu |>
  mutate(edu_inequality_missing_country = as.numeric(is.na(Education_Inequality))) |>
  group_by(Country) |>
  summarize(num_edu_inequality_missing_country = sum(edu_inequality_missing_country))
```

#### 4.2.2 Data missing for each year

For `Education_Level`,

```{r}
d_edu |>
  mutate(edu_level_missing_year = as.numeric(is.na(Education_Level))) |>
  group_by(Year) |>
  summarize(num_edu_level_missing_year = sum(edu_level_missing_year))
```

For `Education_Inequality`,

```{r}
d_edu |>
  mutate(edu_inequality_missing_year = as.numeric(is.na(Education_Inequality))) |>
  group_by(Year) |>
  summarize(num_edu_inequality_missing_year = sum(edu_inequality_missing_year))
```

### 4.2 Country-level indicators

#### 4.2.1 Average level of education quality, 1984-2022

For `Education_Level`:

```{r}

# Countries with highest average education level

d_edu |>
  filter(Year >= 1984 & Year <= 2022) |>
  arrange(Year) |>
  group_by(Country) |>
  summarize(Average_Education_Level = mean(Education_Level, na.rm = TRUE)) |>
  ungroup() |>
  arrange(desc(Average_Education_Level))

# Countries with lowest average education level

d_edu |>
  filter(Year >= 1984 & Year <= 2022) |>
  arrange(Year) |>
  group_by(Country) |>
  summarize(Average_Education_Level = mean(Education_Level, na.rm = TRUE)) |>
  ungroup() |>
  arrange(Average_Education_Level)
```

For `Education_Inequality`:

```{r}
# Countries with least amount of inequality

d_edu |>
  filter(Year >= 1984 & Year <= 2022) |>
  arrange(Year) |>
  group_by(Country) |>
  summarize(Average_Education_Inequality = mean(Education_Inequality, na.rm = TRUE)) |>
  ungroup() |>
  arrange(Average_Education_Inequality)

# Countries with most amount of inequality

d_edu |>
  filter(Year >= 1984 & Year <= 2022) |>
  arrange(Year) |>
  group_by(Country) |>
  summarize(Average_Education_Inequality = mean(Education_Inequality, na.rm = TRUE)) |>
  ungroup() |>
  arrange(desc(Average_Education_Inequality))
```

#### 4.2.2 Change of education quality, 1984-2022

Part 1: Total Change:

For `Education_Level` :

```{r}
# Countries with most increase in average education level

d_edu |>
  filter(Year >= 1984 & Year <= 2022) |>
  arrange(Year) |>
  group_by(Country) |>
  summarise(Education_Level_Change =
              last(Education_Level, na_rm = TRUE) -
              first(Education_Level, na_rm = TRUE)) |>
  ungroup() |>
  arrange(desc(Education_Level_Change))

# Countries with least increase in education years
d_edu |>
  filter(Year >= 1984 & Year <= 2022) |>
  arrange(Year) |>
  group_by(Country) |>
  summarise(Education_Level_Change =
              last(Education_Level, na_rm = TRUE) -
              first(Education_Level, na_rm = TRUE)) |>
  ungroup() |>
  arrange((Education_Level_Change))
```

For `Education_Inequality`:

```{r}
# Countries with most decline of education inequality
d_edu |>
  filter(Year >= 1984 & Year <= 2022) |>
  arrange(Year) |>
  group_by(Country) |>
  summarise(Education_Inequality_Change =
              last(Education_Inequality, na_rm = TRUE) -
              first(Education_Inequality, na_rm = TRUE)) |>
  ungroup() |>
  arrange(desc(Education_Inequality_Change))

# Countries with most increase in education inequality

d_edu |>
  filter(Year >= 1984 & Year <= 2022) |>
  arrange(Year) |>
  group_by(Country) |>
  summarise(Education_Inequality_Change =
              last(Education_Inequality, na_rm = TRUE) -
              first(Education_Inequality, na_rm = TRUE)) |>
  ungroup() |>
  arrange(Education_Inequality_Change)
```

Part 2: Year-on-year change

For `Education_Level`,

```{r}
d_edu |>
  group_by(Country) |>
  arrange(Year) |>
  mutate(edu_level_yoy_change = Education_Level - lag(Education_Level, n = 1)) |>
  ungroup() |>
  select(Country, Year, edu_level_yoy_change) |>
  arrange(Country, Year)
```

For `Education_Inequality` ,

```{r}
d_edu |>
  group_by(Country) |>
  arrange(Year) |>
  mutate(edu_inequality_yoy_change = Education_Inequality - lag(Education_Inequality, n = 1)) |>
  ungroup() |>
  select(Country, Year, edu_inequality_yoy_change) |>
  arrange(Country, Year)
```

### 4.3 Discussion

Which countries perform the best and the worst in terms of education quality in the past four decades?

If we look at the **average performance** over the years, then developed countries such as Germany, Australia, and the UK has the best education quality, where an average citizen would have nearly 13 years of education. Meanwhile, developing countries such as Burkina Faso, Niger, and Mali has the worst education quality, where an average citizen only receives about 1 year of education.

In terms of education inequality, countries such as Austria, Barbados, and Denmark are the most equal, while countries such as Burkina Faso, Mali, and Niger have serious inequality.

However, if we look at **improvements**, then countries such as Botswana, Singapore, and Libya see the largest increase in education years of their citizens, while countries such as Tajikistan, North Korea, and Russia see little increase, or even decline, in this regard.

For education inequality, only 7 countries have improvements: Costa Rica, New Zealand, Spain, Trinidad and Tobago, Switzerland, Lebanon, and Seychelles. In other countries, inequality exacerbated, with Nepal, Botswana, and Haiti having the worse situation.
